---
title: "Reproductible research"
author: "Dmitriy Kairbayev"
date: "01 12 2020"
output: html_document
---
##Assignment Instructions
1.Code for reading in the dataset and/or processing the data
2.Histogram of the total number of steps taken each day
3.Mean and median number of steps taken each day
4.Time series plot of the average number of steps taken
5.The 5-minute interval that, on average, contains the maximum number of steps
6.Code to describe and show a strategy for imputing missing data
7.Histogram of the total number of steps taken each day after missing values are imputed
8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
9.All of the R code needed to reproduce the results (numbers, plots, etc.) in the report

## Including Plots

```{r, echo = TRUE}
activity <- read.csv("activity.csv")
```


```{r, echo = TRUE}
dim(activity)
names(activity)
head(activity)
str(activity)
#total number of missing data
sum(is.na(activity$steps))/dim(activity)[[1]]
#transforming the date column into date format using lubridate
library(lubridate)
activity$date<-ymd(activity$date)
length(unique(activity$date))
```


```{r, echo = TRUE}
library("ggplot2")
activity$date <- as.POSIXct(activity$date, format="%Y-%m-%d")
weekdays <- weekdays(activity$date)
activity <- cbind(activity, weekdays)
summary(activity)
```

## What is mean total number of steps taken per day?

```{r, echo = TRUE}
total_steps <- with(activity, aggregate(steps, by = list(date), FUN = sum, na.rm = TRUE))
names(total_steps) <- c("date", "steps")
hist(total_steps$steps, main = "Total number of steps taken per day", xlab = "Total number of steps taken per day", 
     col = "darkblue", ylim = c(0, 20), breaks = seq(0, 25000))

mean(total_steps$steps)
median(total_steps$steps)
```

## mean = 9354.23
## median = 10395

## What is the average daily activity pattern?

```{r, echo = TRUE}
average_activity <- aggregate(activity$steps, by = list(activity$interval), FUN = mean, na.rm = TRUE)
names(average_activity) <- c("interval", "mean")
plot(average_activity$interval, average_activity$mean, type = "l", col = "darkblue", lwd = 2,
     xlab = "Interval", ylab = "Average number of steps", main = "Average number of steps per intervals")

average_activity[which.max(average_activity$mean), ]$interval
```

## 835


##Imputing missing values

```{r, echo = TRUE}
sum(is.na(activity$steps))
```

## na = 2304

```{r, echo = TRUE}
imputed_steps <- average_activity$mean[match(activity$interval, average_activity$interval)]

activity_imputed <- transform(activity, steps = ifelse(is.na(activity$steps), yes = imputed_steps, no = activity$steps))
total_steps_imputed <- aggregate(steps ~ date, activity_imputed, sum)
names(total_steps_imputed) <- c("date", "daily_steps")

hist(total_steps_imputed$daily_steps, col = "darkblue", xlab = "Total steps per day", ylim = c(0,30), 
     main = "Total number of steps taken each day", breaks = seq(0,25000,by=2500))

mean(total_steps_imputed$daily_steps)
median(total_steps_imputed$daily_steps)

```

## mean = 10766.19
## median = 10766.19

## Are there differences in activity patterns between weekdays and weekends?

```{r, echo = TRUE}
activity$date <- as.Date(strptime(activity$date, format="%Y-%m-%d"))
activity$day <- weekdays(activity$date)
for (i in 1:nrow(activity)) {
    if (activity[i,]$day %in% c("Saturday","Sunday")) {
        activity[i,]$day<-"weekend"
    }
    else{
        activity[i,]$day<-"weekday"
    }
}
stepsByDay <- aggregate(activity$steps ~ activity$interval + activity$day, activity, mean)

names(stepsByDay) <- c("interval", "day", "steps")
library(lattice)
xyplot(steps ~ interval | day, stepsByDay, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")
```

## yes, there are